Manuscript draft

### 2019-07-23 manuscript/manuscript draft.Rmd
updated manuscript draft

To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 8 climate variables. For each possible pairing of climate variables, three models were specified: a null model; a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect. AIC values were calculated for the models and used to select the best model. R-squared values were calculated for the best model. Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.

@@

Including a second climate term in the models didn't tend to significantly increase the explanatory power of the models. The best multivariate models for each carbon flux have r-squared values in the range of those recorded for models with a single climate term. However, the multivariate models do capture the additive and interactive effects of temperature and water availability.

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}", "R\\textsubscript{root}")
kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary. For each carbon flux, correlation coefficients are presented for the two climate variables which explain the most variation when included in a multivariate model. If the model includes an interactive effect, correlation coefficients for the interaction are presented. Results for R\\textsubscript{auto} are not presented here due to the small sample size (n = 14).  ", escape = FALSE, row.names = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  column_spec(1:10, width = "2cm") %>%
  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 6, " "))

@@

| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | Multivariate model |
|----------|----------|----------|----------|----------|----------|
| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 | NA |
| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.74 |
| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.56 |
| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.49 |
| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 | 0.66 |
| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | 0.26 |
| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.36 |
| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.47 |

>>>>

To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 7 climate variables: mean annual temperature; temperature seasonality; mean annual precipitation; precipitation seasonality; potential evapotranspiration; vapour pressure deficit; and solar radiation. For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. In addition a null model, and models including each climate term as a individual (univariate) linear term were specified. As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect. AIC values were calculated for the models and used to select the best model. R-squared values were calculated for the best model. Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.

@@

Including a second climate term in the models tended to have only a small effect on the explanatory power of the models. For ANPP~woody stem~ and R~auto~, including two climate terms did not improve the explantory power of the models: mean annual temperature alone was the best climate predictor for both of these fluxes. For all the remaining carbon fluxes, including a second climate term did improve the explanatory power of the models. ANPP and ANPP~foliage~ were best explained by an additive model, in both cases including mean annual temperature and potential evapotranspiration as the climate terms. GPP, NPP, BNPP, BNPP~fine root~, and R~root~ were all best explained by multivariate models which included an interaction term between the two climate terms. The best models for GPP, NPP, and R~root~ all included a temperature term and a precipitation term, indicating that although precipitation isn't a strong predictor for these variables, the interaction between temperature and water availability is important.

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}","R\\textsubscript{auto}", "R\\textsubscript{root}")
kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary. For each carbon flux, correlation coefficients are presented for the two climate variables which explain the most variation when included in a multivariate model. If the model includes an interactive effect, correlation coefficients for the interaction are presented. For carbon fluxes that were best predicted by a single climate term, correlation coefficients are presented for that term only.  ", escape = FALSE, row.names = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  column_spec(1:10, width = "2cm") %>%
  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 7, " "))

@@

| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | Multivariate model |
|----------|----------|----------|----------|--------------|----------|
| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 | NA |
| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.73 |
| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.57 |
| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.50 |
| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 | 0.66 |
| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | NA |
| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.38 |
| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.43 |
